\ieeetitle

### 2016-12-29 report/abstract.tex
pass on thesis

\begin{abstract}
\epto is one of the recently introduced total order algorithms for large-scale distributed systems and claims to provide total order and scalability at the same time. In this paper, we verify this claim by implementing EpTO \autocite{matos2015epto} and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics. Our results shows \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
\end{abstract}

>>>>

\hm{Sorry for the delay. I am in dependency hell!!! Why on earth did you use tikz pictures? I gave up for now and commented all such figures. Can you compile them separately and include the pdf in the document?}

\begin{abstract}
  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by implementing EpTO, and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics. Our results shows \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
\end{abstract}

### 2016-12-29 report/definitions.tex
pass on thesis

\section{Definitions}
\label{sec:definitions}
A process or peer is defined as an actor in our system running the application that needs total order. Each process will communicate with other processes in the distributed system, exchange events, and order them together.
\par
An event is defined as data sent at a given time by a peer. For example, we could imagine a system where each process publishes some data to other peers. The moment where we publish data combined with the data is called an event.
\par
A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic.

We define \epto scaling well as it is defined in  \autocite{matos2015epto}:  ``The number
of messages transmitted per process per round is logarithmic
in the number of processes, ...''. The number of rounds needed to deliver is logarithmic as well.
\par
Since \epto uses a probabilistic agreement instead of a deterministic agreement, there is a very low arbitrary probability a peer does not receive an event. In this instance there will be a hole in the sequence of delivered events but even in this case, the order of the delivered events will be protected by \epto's deterministic ordering algorithm, thus the total order property is preserved.
\par
We write the cluster parameters as $(n,e)$, where $n$ designates the total number of peers and $e$ denominates the global event throughput per second. We tested \epto and \jgroups with three different cluster parameters:
\begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
	\item[\textbf{$(50,50)$}:] 50 peers with a global event throughput of 50 events per second.
	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.
\end{description}

>>>>

\section{Definitions}
\label{sec:definitions}
A process or peer is defined as an actor in our system running the application that needs total order. Each process communicates with other processes in the distributed system, exchange events, and order them together. \hm{I changed it, but you switched from the present to the future tense in the same sentence. You might as well check the rest of the document for other instances, if any.}
\par
An event is defined as data sent at a given time by a peer. For example, we could imagine a system where each process publishes some data to other peers. The moment where we publish data combined with the data is called an event. \hm{Not clear. Define it mathematically with something like $e_k=(id,time,payload)$ where the time is the (local) time at the process when the event is created.}
\par
A ball is a set of events bundled together and sent as one package. We use balls in \epto to reduce network traffic. \hm{Not clear. A ball is an abstraction. We reduce traffic by bundling events.}

We define \epto scaling well as it is defined in  \autocite{matos2015epto}:  ``The number
of messages transmitted per process per round is logarithmic
in the number of processes, ...''. The number of rounds needed to deliver is logarithmic as well. \hm{This paragraph is confusing. Is it necessary?}
\par
Since \epto uses a probabilistic agreement instead of a deterministic agreement, there is a nonzero probability that a peer does not receive an event. In this instance there will be a hole in the sequence of delivered events, but the order of the delivered events is protected by \epto's deterministic ordering algorithm, thus the total order property is preserved. The probability that a peer does not receive an event is controlled by the \epto and depends on various parameters such as the fan-out, the number of rounds (time to live), the number of nodes, and the network conditions. In particular, this probability can be made smaller than the probability of catastrophic network failure. \hm{I rewrote this paragraph. Please reread.}
\par
We write the cluster parameters as $(n,e)$, where $n$ designates the total number of peers and $e$ denominates the global event throughput per second. We tested \epto and \jgroups with three different cluster parameters:
\begin{description}[\IEEEsetlabelwidth{$(100,100)$:}]
	\item[\textbf{$(50,50)$}:] 50 peers with a global event throughput of 50 events per second.
	\item[\textbf{$(50,100)$}:] 50 peers with a global event throughput of 100 events per second.
	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.
\end{description}

### 2016-12-29 report/evaluation.tex
pass on thesis

\section{Evaluation}
\label{sec:evaluation}
\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
In this section we present our results. In the following benchmarks \epto uses a constant $c$ to set the probability of a hole appearing. We set $c = 4$ in all our benchmarks so as to not experience any holes as \jgroups does not produce holes under normal conditions. We use a $\delta$ period of \SI{100}{\milli\second} for tests with no churn or synthethic churn and \SI{250}{\milli\second} for tests following a real trace.
\par 
We used specific settings recommended by \jgroups for a big cluster.
\par
Every \jgroups test run with churn is run once killing the coordinator and once not killing it. We use the following nomenclature to differentiate both tests:
\begin{description}[\IEEEsetlabelwidth{\jgroups-nocoord: }]
	\item[\textbf{\jgroups-coord}:] The coordinator is killed once.
	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive throughout the experiment.
\end{description}
\par
\textbf{Testbed.} \jt{Most of this paragraph is copied from ErasureBench as we use the same cluster and Docker as well. Is this fine as is?}Our cluster consists of a cluster of 20 machines interconnected by a \SI{1}{\gbps} switched network. Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM. We deploy 12 virtual machines on these hosts. Each virtual machine has acces to 4 CPUs and 4 VCPUs as well as \SI{7}{\giga\byte} of RAM. We use KVM as our hypervisor. Each VM uses Debian as its O/S. EpTOTester is packaged as docker images. We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services. Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.
\par
\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}. When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}. The benchmarks following a real trace are run 5 times. The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}. Every benchmark run with churn is run with the $(100,50)$ parameters. When we have a global even throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput. For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.
\par
We compared \epto and \jgroups based on their bandwidth, the time to deliver every expected events, their local dissemination stretch and finally the number of events sent.
\subsection{Bandwidth}
The initial bandwidth peaks observed for \epto are due to the PSS initialization. When a node starts, it received an initial view and then quickly run the PSS four times to obtain a random enough view.
\begin{figure}[hpt]
	\centering
	\input{figures/no-churn/bandwidth-figure.tex}
	\vspace{-2mm} 
	\caption{Bandwidth percentiles of a node during an experiment}
	\vspace{-2mm} 
	\label{fig:bandwidth}
\end{figure}
In \autoref{fig:bandwidth} we observe that \epto has a worse baseline compared to \jgroups. It uses a median bandwidth of approximately \SI{1}{\mbps} for $(50,50)$ whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}. However, in \jgroups most of the work is done solely by the coordinator. We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.6}{\mbps}.

@@

\begin{figure}[hpt]
	\centering
	\input{figures/synthetic-churn/bandwidth-figure-churn.tex}
	\vspace{-2mm} 
	\caption{Bandwidth percentiles of a node during an experiment with churn}
	\vspace{-2mm} 
	\label{fig:bandwidth-churn}
\end{figure}

@@

%\begin{figure}[h]
%	\centering
%	\input{figures/synthetic-churn/total-bytes-churn.tex}
%	\vspace{-2mm} 
%	\caption{Total bytes sent/received during an average experiment with churn}
%	\vspace{-2mm} 
%	\label{fig:total-bandwidth-churn}
%\end{figure}
In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn. One hypothesis for this is that a JGroups replica takes a longer time to start up compared to stopping a replica. Therefore the overall benchmark has a longer time with less than 100 replicas. We also do not see a difference whether we kill the coordinator or not. This can be explained by the fact that before the detection of the faulty coordinator \jgroups is forced to a stop for time up to \SI{20}{\second}. The big spike afterwards compensates for this hole.
\subsection{Local Times}
\label{sub:local-times}
\begin{figure*}[hpt]
	\centering
	\input{figures/no-churn/local-times.tex}
	\vspace{-2mm} 
	\caption{Local dissemination times}
	\vspace{-2mm}
	\label{fig:local-times} 
\end{figure*}

\begin{figure}[hpt]
	\centering
	\input{figures/synthetic-churn/local-times-churn.tex}
	\vspace{-2mm} 
	\caption{Local dissemination times with churn}
	\vspace{-2mm} 
	\label{fig:local-times-churn} 
\end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}. We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system.
\subsection{Global Times}
\begin{figure*}[hpt]
	\centering
	\input{figures/no-churn/global-times.tex}
	\vspace{-2mm} 
	\caption{Global dissemination times}
	\vspace{-2mm}
	\label{fig:global-times}  
\end{figure*}

\begin{figure}[hpt]
	\centering
	\input{figures/synthetic-churn/global-times-churn.tex}
	\vspace{-2mm} 
	\caption{Global dissemination times with churn}
	\vspace{-2mm} 
	\label{fig:global-times-churn} 
\end{figure}
We computed global times as well. They are represented in \autoref{fig:global-times} and \autoref{fig:global-times-churn}. These global times are of less interest than their local counterpart as the differences in clocks between hosts can skew this measurement.

Nonetheless, here too we can see that \epto is consistently slower than \jgroups for the same reason as stated in \autoref{sub:local-times}.
\subsection{Local Dissemination stretch}
\begin{figure}[hpt]
	\centering
	\input{figures/no-churn/local-delta.tex}
	\vspace{-2mm} 
	\caption{Local dissemination stretch}
	\vspace{-2mm}
	\label{fig:local-delta}  
\end{figure}
In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch. The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.
\par
\jgroups is usually much faster than \epto in a perfect environment. This is expected as the benchmarks involve a small number of nodes and are performed in a LAN environment with minimal latency. The median dissemination stretch of \jgroups is around \SI{7}{\milli\second} where as the median dissemination stretch of \epto is around \SI{630}{\milli\second} for $(100,50)$. When increasing the number of peers, \jgroups starts to have longer delivery times for some outliers. A small portion of the local dissemination stretches are really fast (\SI{1}{\milli\second} or lower). This happens when the coordinator itself sends an event. Since it has to deliver it to himself the local dissemination stretch is intra-process and thus extremely fast.

\begin{figure}[hpt]
	\centering
	\input{figures/synthetic-churn/local-delta-churn.tex}
	\vspace{-2mm} 
	\caption{Local dissemination stretch with churn}
	\vspace{-2mm}
	\label{fig:local-delta-churn}   
\end{figure}
In \autoref{fig:local-delta-churn} We see a completely different picture. When under churn, the 95th percentile of \jgroups is at \SI{31}{\milli\second} compared to \SI{14}{\milli\second} when there is no churn. The highest percentiles are at more than \SI{10}{\second}. This effect is due to the coordinator dying as we clearly see that it does not happen when we do not kill it.

@@

\begin{figure}[hpt]
	\input{figures/real-churn/bandwidth-figure-real-churn.tex}
	\vspace{-2mm} 
	\caption{Bandwidth percentiles of a node during an experiment with a real churn}
	\vspace{-2mm} 
	\label{fig:bandwidth-real-churn}
\end{figure}

@@

\begin{figure}[hpt]
	\input{figures/real-churn/local-times-real-churn.tex}
	\vspace{-2mm} 
	\caption{Local dissemination times}
	\vspace{-2mm}
	\label{fig:local-times-real-churn} 
\end{figure}
\begin{figure}[hpt]
	\input{figures/real-churn/global-times-real-churn.tex}
	\vspace{-2mm} 
	\caption{Global dissemination times with a real churn}
	\vspace{-2mm} 
	\label{fig:global-times-real-churn} 
\end{figure}

@@

\begin{figure}[hpt]
	\input{figures/real-churn/local-delta-real-churn.tex}
	\vspace{-2mm} 
	\caption{Local dissemination stretch with a real churn}
	\vspace{-2mm}
	\label{fig:local-delta-real-churn}   
\end{figure}

>>>>

\section{Evaluation}
\label{sec:evaluation}
%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
%\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
%\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
%\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
In this section we present our results. In the following benchmarks \epto uses a constant $c$ to set the probability of a hole appearing. We set $c = 4$ in all our benchmarks so as to not experience any holes as \jgroups does not produce holes under normal conditions. We use a $\delta$ period of \SI{100}{\milli\second} for tests with no churn or synthethic churn and \SI{250}{\milli\second} for tests following a real trace.
\par 
We used specific settings recommended by \jgroups for a big cluster.
\par
Every \jgroups test run with churn is run once killing the coordinator and once not killing it. We use the following nomenclature to differentiate both tests:
\begin{description}[\IEEEsetlabelwidth{\jgroups-nocoord: }]
	\item[\textbf{\jgroups-coord}:] The coordinator is killed once.
	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive throughout the experiment.
\end{description}
\par
\textbf{Testbed.} \jt{Most of this paragraph is copied from ErasureBench as we use the same cluster and Docker as well. Is this fine as is?}Our cluster consists of a cluster of 20 machines interconnected by a \SI{1}{\gbps} switched network. Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM. We deploy 12 virtual machines on these hosts. Each virtual machine has acces to 4 CPUs and 4 VCPUs as well as \SI{7}{\giga\byte} of RAM. We use KVM as our hypervisor. Each VM uses Debian as its O/S. EpTOTester is packaged as docker images. We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services. Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.
\par
\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}. When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}. The benchmarks following a real trace are run 5 times. The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}. Every benchmark run with churn is run with the $(100,50)$ parameters. When we have a global even throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput. For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.
\par
We compared \epto and \jgroups based on their bandwidth, the time to deliver every expected events, their local dissemination stretch and finally the number of events sent.
\subsection{Bandwidth}
The initial bandwidth peaks observed for \epto are due to the PSS initialization. When a node starts, it received an initial view and then quickly run the PSS four times to obtain a random enough view.
% \begin{figure}[hpt]
% 	\centering
% 	\input{figures/no-churn/bandwidth-figure.tex}
% 	\vspace{-2mm} 
% 	\caption{Bandwidth percentiles of a node during an experiment}
% 	\vspace{-2mm} 
% 	\label{fig:bandwidth}
% \end{figure}
In \autoref{fig:bandwidth} we observe that \epto has a worse baseline compared to \jgroups. It uses a median bandwidth of approximately \SI{1}{\mbps} for $(50,50)$ whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}. However, in \jgroups most of the work is done solely by the coordinator. We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.6}{\mbps}.

@@

% \begin{figure}[hpt]
% 	\centering
% 	\input{figures/synthetic-churn/bandwidth-figure-churn.tex}
% 	\vspace{-2mm} 
% 	\caption{Bandwidth percentiles of a node during an experiment with churn}
% 	\vspace{-2mm} 
% 	\label{fig:bandwidth-churn}
% \end{figure}

@@

%\begin{figure}[h]
%	\centering
%	\input{figures/synthetic-churn/total-bytes-churn.tex}
%	\vspace{-2mm} 
%	\caption{Total bytes sent/received during an average experiment with churn}
%	\vspace{-2mm} 
%	\label{fig:total-bandwidth-churn}
%\end{figure}
In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn. One hypothesis for this is that a JGroups replica takes a longer time to start up compared to stopping a replica. Therefore the overall benchmark has a longer time with less than 100 replicas. We also do not see a difference whether we kill the coordinator or not. This can be explained by the fact that before the detection of the faulty coordinator \jgroups is forced to a stop for time up to \SI{20}{\second}. The big spike afterwards compensates for this hole.
\subsection{Local Times}
\label{sub:local-times}
% \begin{figure*}[hpt]
% 	\centering
% 	\input{figures/no-churn/local-times.tex}
% 	\vspace{-2mm} 
% 	\caption{Local dissemination times}
% 	\vspace{-2mm}
% 	\label{fig:local-times} 
% \end{figure*}

% \begin{figure}[hpt]
% 	\centering
% 	\input{figures/synthetic-churn/local-times-churn.tex}
% 	\vspace{-2mm} 
% 	\caption{Local dissemination times with churn}
% 	\vspace{-2mm} 
% 	\label{fig:local-times-churn} 
% \end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}. We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system.
\subsection{Global Times}
% \begin{figure*}[hpt]
% 	\centering
% 	\input{figures/no-churn/global-times.tex}
% 	\vspace{-2mm} 
% 	\caption{Global dissemination times}
% 	\vspace{-2mm}
% 	\label{fig:global-timexss}  
% \end{figure*}

% \begin{figure}[hpt]
% 	\centering
% 	\input{figures/synthetic-churn/global-times-churn.tex}
% 	\vspace{-2mm} 
% 	\caption{Global dissemination times with churn}
% 	\vspace{-2mm} 
% 	\label{fig:global-times-churn} 
% \end{figure}
We computed global times as well. They are represented in \autoref{fig:global-times} and \autoref{fig:global-times-churn}. These global times are of less interest than their local counterpart as the differences in clocks between hosts can skew this measurement.

Nonetheless, here too we can see that \epto is consistently slower than \jgroups for the same reason as stated in \autoref{sub:local-times}.
\subsection{Local Dissemination stretch}
% \begin{figure}[hpt]
% 	\centering
% 	\input{figures/no-churn/local-delta.tex}
% 	\vspace{-2mm} 
% 	\caption{Local dissemination stretch}
% 	\vspace{-2mm}
% 	\label{fig:local-delta}  
% \end{figure}
In \autoref{fig:local-delta}, We see the percentiles of the local dissemination stretch. The local dissemination stretch is the time measurement between the sending of an event by a peer and the delivery of this event locally.
\par
\jgroups is usually much faster than \epto in a perfect environment. This is expected as the benchmarks involve a small number of nodes and are performed in a LAN environment with minimal latency. The median dissemination stretch of \jgroups is around \SI{7}{\milli\second} where as the median dissemination stretch of \epto is around \SI{630}{\milli\second} for $(100,50)$. When increasing the number of peers, \jgroups starts to have longer delivery times for some outliers. A small portion of the local dissemination stretches are really fast (\SI{1}{\milli\second} or lower). This happens when the coordinator itself sends an event. Since it has to deliver it to himself the local dissemination stretch is intra-process and thus extremely fast.

% \begin{figure}[hpt]
% 	\centering
% 	\input{figures/synthetic-churn/local-delta-churn.tex}
% 	\vspace{-2mm} 
% 	\caption{Local dissemination stretch with churn}
% 	\vspace{-2mm}
% 	\label{fig:local-delta-churn}   
% \end{figure}
In \autoref{fig:local-delta-churn} We see a completely different picture. When under churn, the 95th percentile of \jgroups is at \SI{31}{\milli\second} compared to \SI{14}{\milli\second} when there is no churn. The highest percentiles are at more than \SI{10}{\second}. This effect is due to the coordinator dying as we clearly see that it does not happen when we do not kill it.

@@

% \begin{figure}[hpt]
% 	\input{figures/real-churn/bandwidth-figure-real-churn.tex}
% 	\caption{Bandwidth percentiles of a node during an experiment with a real churn}
% 	\label{fig:bandwidth-real-churn}
% \end{figure}

@@

% %\end{figure}
% \begin{figure}[hpt]
% 	\input{figures/real-churn/local-times-real-churn.tex}
% 	\caption{Local dissemination times}
% 	\label{fig:local-times-real-churn} 
% \end{figure}
% \begin{figure}[hpt]
% 	\input{figures/real-churn/global-times-real-churn.tex}
% 	\caption{Global dissemination times with a real churn}
% 	\label{fig:global-times-real-churn} 
% \end{figure}

@@

% \begin{figure}[hpt]
% 	\input{figures/real-churn/local-delta-real-churn.tex}
% 	\caption{Local dissemination stretch with a real churn}
% 	\label{fig:local-delta-real-churn}   
% \end{figure}

### 2016-12-29 report/framework.tex
WIP report
pass on thesis

\section{EpToTester}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
\subsection{Architecture}
\autoref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can \epto-broadcast and \epto-deliver events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL). They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not required. In practice it works well, but using a DHT is certainly a possibility.
\subsection{Implementation}
We implement the data sent as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port. We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST. Finally, EpToTester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
\subsection{Deployment}
\begin{figure}[htp]
	\centering
	\input{figures/architecture.tex}
	\vspace{-2mm} 
	\caption{EpTOTester architecture}
	\vspace{-2mm} 
	\label{fig:complete-architecture}
\end{figure}
To deploy \epto effectively we need two different Docker Swarm services. One containing the tracker and one containing all \epto replicas. Both of these services use the same network overlay to communicate. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications. Every benchmark is started through a Python script available on the master mode. This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle is used to automate the image building in the python scripts. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}. \jt{The figure is inspired from the one in ErasureBench. How should I indicate this?}
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}. The synthetic churn provided by it was improved to support adding and removing nodes at the same time.

>>>>

\section{EpToTester}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation in written with benchmarking in mind, the code could be adapted for real applications with only minimal changes to the sources. \hm{Write that the code is publicly available at some point, and where to find it.}
\subsection{Architecture}
\autoref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can \epto-broadcast and \epto-deliver events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL). They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not required. In practice it works well, but using a DHT is certainly a possibility.
\subsection{Implementation}
We implement the data sent as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port. \hm{jgroups appears from nowhere. Rewrite that is used to compare with EpTO.} We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST. Finally, EpToTester \hm{The name is not capitalized properly (o instead of O). Write as a latex definition like epto, and change everywhere in the document.} is open-sourced on Github \hm{There it is. This should have been said earlier, when EpTOTester is first introduced.}\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
\subsection{Deployment}
% \begin{figure}[htp]
% 	\centering
% 	\input{figures/architecture.tex}
% 	\vspace{-2mm} 
% 	\caption{EpTOTester architecture}
% 	\vspace{-2mm} 
% 	\label{fig:complete-architecture}
% \end{figure}
To deploy \epto effectively we need two different Docker Swarm services: one containing the tracker and one containing all \epto replicas. Both of these services use the same network overlay to communicate. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications. Every benchmark is started through a Python script available on the master mode.\hm{What is the master mode? Node maybe?} This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle \hm{Never heard of this. Please add reference or url in a footnote.} is used to automate the image building in the python scripts. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}. \jt{\autoref{fig:complete-architecture} is inspired from the one in ErasureBench. How should I indicate this?}\hm{Discussed during our call. A citation to Sebastien's paper in a footnote will do.}
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}. The synthetic churn provided by it was improved to support adding and removing nodes at the same time.
